Exploring the Soil Microbiome coupling with the Crete island Data Cube

### 2024-04-06 manuscript.tex
update keywords

\abstract{
    Microbes are known for their versatility, abundance
    and influence on soil ecosystem functioning.
    A synthesized knowledge base of microbial biodiversity, in terms of
    ecological and remote-sensing data remains a major challenge.
    Many worldwide studies have been published regarding soil
    microbiome ecosystems, though there are still many blind spots.
    Islands can be important case studies for this integration for more resolute and dense samplings.
    Here, we utilize the Island Sampling Day Crete 2016 microbial 16S rRNA gene
    amplicon data, integrated with soil and remote
    sensing data, to decipher the drivers of ecosystem function of the island.
    The Island Sampling Day Crete 2016 project has collected 144 topsoil samples
    from 72 sites, capturing a lot of this diversity, accompanied by FAIR
    (Findable, Accessible, Interoperable and Reproducible) data by design. 
    Cretan macroecology has been studied for centuries for its diverse  and endemic
    fauna and flora.
    In addition, Crete has been considered as a miniature continent with high contrasts in
    vegetation cover, elevation, climatic conditions. 
    We show that, higher altitudes in Crete found to
    be inhabited by a more diverse number of microorganisms, a pattern commonly
    seen in several faunistic groups, such as arthropods.
    The integration of the spatial data presented here enabled warning signals
    in pristine and grazing ecosystems.
    These results along with the
    climatic and desertification index influences on the soil microbiome of Crete,
    provide the basis to identify major drivers of biodiversity, to evaluate hotspots
    and contribute to foreknowledge of threatened ecosystems.
}

\keywords{soil microbiome, networks, island biogeography, Crete, data integration, FAIR data}

>>>>

\abstract{
    Microbes are known for their versatility, abundance
    and influence on soil ecosystem functioning.
    A synthesized knowledge base of microbial biodiversity, in terms of
    ecological and remote-sensing data remains a major challenge.
    Many worldwide studies have been published regarding soil
    microbiome ecosystems, though there are still many blind spots.
    Islands can be important case studies for this integration for more resolute and dense samplings.
    Here, we utilize the Island Sampling Day Crete 2016 microbial 16S rRNA gene
    amplicon data, integrated with soil and remote
    sensing data, to decipher the drivers of ecosystem function of the island.
    The Island Sampling Day Crete 2016 project has collected 144 topsoil samples
    from 72 sites, capturing a lot of this diversity, accompanied by FAIR
    (Findable, Accessible, Interoperable and Reproducible) data by design. 
    Cretan macroecology has been studied for centuries for its diverse  and endemic
    fauna and flora.
    In addition, Crete has been considered as a miniature continent with high contrasts in
    vegetation cover, elevation, climatic conditions. 
    We show that, higher altitudes in Crete found to
    be inhabited by a more diverse number of microorganisms, a pattern commonly
    seen in several faunistic groups, such as arthropods.
    The integration of the spatial data with state of the art methods enabled warning signals
    in pristine and grazing ecosystems.
    These results along with the
    climatic and desertification index influences on the soil microbiome of Crete,
    provide the basis to identify major drivers of biodiversity, to evaluate hotspots
    and contribute to foreknowledge of threatened ecosystems.
}

\keywords{soil microbiome, networks, island biogeography, Crete, spatial data, remote sensing, data integration, FAIR data}
